---
title: "Stroke Prediction"
output:
  flexdashboard::flex_dashboard:
    theme:
      bg: '#101010'
      fg: '#FDF7F7'
      primary: '#D6D5D5'
      base_font:
        google: Prompt
      code_font:
        google: Montserrat
    orientation: columns
    vertical_layout: fill
    source_code: embed
    logo: stroke7.png
    favicon: stroke7.png
runtime: shiny
resource_files:
- low_volume_MLs.R
- XGboost_Final.R
- RanFor_Final.R
- Final_Sina.R
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(readr)
library(vip)
library(janitor)
library(forcats)
library(tidymodels)
library(tidyverse)
library(modelr)
library(readxl)
library(broom)
library(ggfortify)
library(performance)
library(ggplot2)
library(plotly)
library(dplyr)
library(skimr)
library(rsample)
library(themis)
library(yardstick)
library(xgboost)
library(bslib)
library(shinydashboard)
library(shinythemes)
library(knitr)
library(stringr)
library(shinyjs)
library(httr)
library(jsonlite)
library(rsconnect)


# Reactive value to store prediction probability
stroke_risk <- reactiveVal(0)

# Import data
dataRaw <- read_csv("healthcare.csv")
dataClean <- dataRaw %>%
  clean_names() %>%
  mutate(
    ever_married    = if_else(ever_married == "Yes", 1L, 0L),
    residence_type  = if_else(residence_type == "Urban", 1L, 0L),
    gender          = if_else(gender == "Male", 1L, 0L),
    stroke          = factor(stroke, levels = c(0, 1), labels = c("No", "Yes")),
    work_type       = as.factor(work_type),
    smoking_status  = as.factor(smoking_status),
    bmi             = as.numeric(na_if(bmi, "N/A"))
  )

dataClean2 <- dataClean %>%
  drop_na() %>%
  filter(smoking_status != "Unknown")
```


Overview {data-width=800}
===========================================================

Input your Data Here {.sidebar}
-----------------------------------------------------------
##### Predictor's Summary:

```{r side_bar}

selectInput("hist_var", "Select a variable:", 
  choices = c(
    "Age" = "age",
    "BMI" = "bmi",
    "Glucose Level" = "avg_glucose_level",
    "Heart Disease" = "heart_disease",
    "Hypertension" = "hypertension",
    "Smoking Status" = "smoking_status",
    "Ever Married" = "ever_married"
  )
)
```


```{r data_source}
div(style = "
     position: absolute;
     bottom: 20px;
     width: 90%; 
     left: 5%; 
     text-align: center;",
  
  tags$h5("Access Stroke Data", style = "color: #FDF7F7; font-weight: bold; margin-bottom: 10px;"),

  HTML('<a href="https://www.kaggle.com/datasets/fedesoriano/stroke-prediction-dataset" 
           target="_blank" 
           class="btn btn-secondary" 
           role="button" 
           style="width: 100%;">
           Go to Data Source
        </a>')
)
```


Column {data-width=500}
-----------------------------------------------------------

### **Project Description**

<p>According to the World Health Organization (WHO) stroke is the 2nd leading cause of death globally, responsible for approximately 11% of total deaths.

A stroke is a medical emergency that occurs when blood flow to a part of the brain is blocked or interrupted, preventing brain cells from getting oxygen and nutrients. Brain cells begin to die within minutes, which can cause permanent damage.

This data set is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, history of cardiovascular diseases, and smoking status. Each row in the data provides relevant information about the patient.

Created by Sina Mahdiani | Data Science 2 final project</p>

### **Data Dictionary for Selected Variables**

```{r Dictionary, echo=FALSE}
data.frame(
  Variables = c("id","gender", "age", "hypertension", "heart_disease", 
                "ever_married", "work_type", "residence_type", 
                "avg_glucose_level", "bmi", "smoking_status", "stroke"),
  Description = c("Unique Identifier",
                  "Binary variable: 0 = Female, 1 = Male", 
                  "Continuous variable — patient’s age in years", 
                  "Binary variable: 0 = No hypertension, 1 = Yes",
                  "Binary variable: 0 = No history of heart disease, 1 = Yes",
                  "Binary variable: 0 = Never married, 1 = Ever married",
                  "Categorical: types of employment (Private, Govt job, etc.)",
                  "Binary: 0 = Rural, 1 = Urban",
                  "Continuous variable — average blood glucose level (mg/dL)",
                  "Continuous variable — Body Mass Index", 
                  "Categorical: never smoked, formerly smoked, smokes",
                  "Whether the patient has had a stroke (Yes or No)"),
  Type = c("Identifier", "Predictor", "Predictor", "Predictor", 
           "Predictor", "Predictor", "Predictor", "Predictor", 
           "Predictor", "Predictor", "Predictor", "Outcome")
) %>% 
  kable()
```


Column {data-width=600}
-----------------------------------------------------------

### **Preview of Dataset**

```{r dataset}
DT::datatable(
  dataClean2,
  options = list(pageLength = 5, scrollX = TRUE),
  rownames = FALSE
)
```

### **Dynamic Variable Distribution Plot**

```{r}
plotlyOutput("hist_plot")
```

```{r}
output$hist_plot <- renderPlotly({
  req(input$hist_var)
  var <- input$hist_var
  values <- dataClean2[[var]]
  is_numeric <- is.numeric(values)
  is_binary <- is_numeric && length(unique(na.omit(values))) == 2

  p <- if (is_binary) {
    ggplot(dataClean2, aes(x = factor(.data[[var]]))) +
      geom_bar(fill = "darkred", color = "#101010", alpha = 0.85)
  } else if (is_numeric) {
    ggplot(dataClean2, aes(x = .data[[var]])) +
      geom_histogram(bins = 30, fill = "orange", color = "#101010", alpha = 0.85)
  } else {
    ggplot(dataClean2, aes(x = .data[[var]])) +
      geom_bar(fill = "darkgreen", color = "#101010", alpha = 0.85)
  }

  p <- p +
    geom_hline(yintercept = 0, color = "white", linewidth = 0.3) +
    geom_vline(xintercept = 0, color = "white", linewidth = 0.3) +
    labs(
      title = paste("Distribution of", str_to_title(gsub("_", " ", var))),
      x = str_to_title(gsub("_", " ", var)),
      y = "Count"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "white"),
      axis.title = element_text(face = "bold", size = 12, color = "white"),
      axis.text = element_text(size = 10, color = "white"),
      legend.text = element_text(color = "white"),
      legend.title = element_text(color = "white"),
      strip.text = element_text(color = "white"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = "#101010", color = NA),
      panel.background = element_rect(fill = "#101010", color = NA)
    )

  ggplotly(p, tooltip = c("x", "y"))
})
```


Classification Results {data-width=800}
===========================================================
```{r models}

#DATA SPLITTING AND RESAMPLING

set.seed(123)
splits <- initial_split(dataClean2, strata = stroke)

dataClean_other <- training(splits)
dataClean_test  <- testing(splits)

# Validation set from training
set.seed(234)
val_set <- validation_split(dataClean_other, strata = stroke, prop = 0.8)

#Recipe
lr_recipe <- recipe(stroke ~ ., data = dataClean_other) %>%
  update_role(id, new_role = "ID") %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors())

# Load pre-trained workflows
lr_workflow_best <- readRDS("lr_model.rds")
rf_workflow_best <- readRDS("rf_model.rds")
xgb_workflow_best <- readRDS("xgb_model.rds")

# Refit models to training/test split
lr_last_fit <- lr_workflow_best %>% last_fit(splits)
rf_last_fit <- rf_workflow_best %>% last_fit(splits)
xgb_last_fit <- xgb_workflow_best %>% last_fit(splits)

```

Sidebar {.sidebar}
-----------------------------------------------------------
#####

```{r ModelChoice}
selectInput("model_choice", "Choose Model:",
            choices = c("Logistic Regression", "Random Forest", "XGBoost"))
```

ROC Curve {data-width=400}
-----------------------------------------------------------

### ROC Curve

```{r ROC}
renderPlot({
  fit <- switch(input$model_choice,
                "Logistic Regression" = lr_last_fit,
                "Random Forest" = rf_last_fit,
                "XGBoost" = xgb_last_fit)

  preds <- fit %>% collect_predictions()
  roc_curve(preds, truth = stroke, .pred_Yes) %>% 
    autoplot() + theme_minimal()
})
```


### Confusion Matrix

```{r ConfMatrix}
plotOutput("conf_matrix_plot", height = "400px")
```

```{r ConfusionMatrix, context="server"}
output$conf_matrix_plot <- renderPlot({
  fit <- switch(input$model_choice,
                "Logistic Regression" = lr_last_fit,
                "Random Forest" = rf_last_fit,
                "XGBoost" = xgb_last_fit)

  preds <- fit %>% collect_predictions()

  cm <- preds %>%
    conf_mat(truth = stroke, estimate = .pred_class)

  autoplot(cm, type = "heatmap") +
    scale_fill_gradient(low = "lightblue", high = "darkred") +
    labs(title = paste("Confusion Matrix -", input$model_choice),
         fill = "Count") +
    theme_minimal(base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "white"),
      axis.title = element_text(size = 12, color = "white"),
      axis.text = element_text(size = 10, color = "white"),
      plot.background = element_rect(fill = "#101010", color = NA),
      panel.background = element_rect(fill = "#101010", color = NA),
      legend.title = element_text(color = "white"),
      legend.text = element_text(color = "white")
    )
})
```

Variable Importance Plot {data-width=400}
-----------------------------------------------------------
### Variable Importance Plot

```{r VIP}

renderPlot({
  fit <- switch(input$model_choice,
                "Logistic Regression" = lr_last_fit,
                "Random Forest" = rf_last_fit,
                "XGBoost" = xgb_last_fit)

  model <- fit %>% extract_fit_parsnip() %>% pluck("fit")

  #Assign different color per model
  color <- switch(input$model_choice,
                  "Logistic Regression" = "orange", 
                  "Random Forest" = "darkgreen",      
                  "XGBoost" = "darkred")              

  vip(model, num_features = 10, 
      aesthetics = list(fill = color, color = color)) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "white"),
      axis.title = element_text(size = 14, face = "bold", color = "white"),
      axis.text = element_text(size = 14, face = "bold", color = "white"),
      plot.background = element_rect(fill = "#101010", color = NA),
      panel.background = element_rect(fill = "#101010", color = NA),
 panel.grid.major = element_line(color = "grey30", size = 0.4),
panel.grid.minor = element_line(color = "grey20", size = 0.2),
    )
})

```

### Metrics

```{r metrics_button}
selectInput("metric_choice", "Choose Metric:",
            choices = c("Accuracy", "Sensitivity", "Specificity", "AUC"))

```

```{r metrics}
renderGauge({
  fit <- switch(input$model_choice,
                "Logistic Regression" = lr_last_fit,
                "Random Forest" = rf_last_fit,
                "XGBoost" = xgb_last_fit)

  preds <- fit %>% collect_predictions()

  metric_value <- switch(input$metric_choice,
                         "Accuracy" = accuracy(preds, truth = stroke, estimate = .pred_class)$.estimate,
                         "Sensitivity" = sens(preds, truth = stroke, estimate = .pred_class)$.estimate,
                         "Specificity" = specificity(preds, truth = stroke, estimate = .pred_class)$.estimate,
                         "AUC" = roc_auc(preds, truth = stroke, .pred_Yes)$.estimate)

  gauge(round(metric_value * 100), min = 0, max = 100, symbol = '%',
        gaugeSectors(success = c(85, 100),
                     warning = c(70, 85),
                     danger = c(0, 70)))
})
```


Calculator {data-width=800}
===========================================================

Sidebar {.sidebar}
-----------------------------------------------------------

#### Parameters:

```{r side_bar2}
sliderInput("age", "Age:", min = 0, max = 100, value = 50)

selectInput("gender", "Gender:", choices = c("Male" = 1, "Female" = 0))

selectInput("smoking", "Smoking Status:", 
            choices = levels(dataClean$smoking_status))

selectInput("work", "Work Type:", choices = levels(dataClean$work_type))

selectInput("residence", "Residence Type:", choices = c("Urban" = 1, "Rural" = 0))

sliderInput("glucose", "Avg Glucose Level:", min = 50, max = 300, value = 100)

sliderInput("bmi", "BMI:", min = 10, max = 60, value = 25)

checkboxInput("married", "Ever Married?", TRUE)

actionButton("predict", "Predict Stroke Risk")
```

Prediction result {data-width=1000}
-----------------------------------------------------------

### Let me guess!

```{r prediction}

observeEvent(input$predict, {
  new_data <- tibble(
    id = 9999,
    age = input$age,
    avg_glucose_level = input$glucose,
    bmi = input$bmi,
    gender = as.integer(input$gender),
    smoking_status = factor(input$smoking, levels = levels(dataClean$smoking_status)),
    work_type = factor(input$work, levels = levels(dataClean$work_type)),
    residence_type = as.integer(input$residence),
    ever_married = as.integer(input$married),
    hypertension = 0,
    heart_disease = 0
  )

  model <- xgb_workflow_best %>% fit(dataClean_other)
  prediction <- predict(model, new_data, type = "prob")
  prob <- round(prediction$.pred_Yes * 100, 1)

  stroke_risk(prob)
})

output$prediction_gauge <- renderGauge({
  prob <- stroke_risk()

  gauge(prob,
        min = 0, max = 100, symbol = '%',
        gaugeSectors(success = c(0, 20), warning = c(20, 50), danger = c(50, 100))
  )
})
```

```{r text-gauge}
output$prediction_text <- renderUI({
  prob <- stroke_risk()

  message <- case_when(
    prob >= 80 ~ "Danger zone! cancel your bungee jump.",
    prob >= 50 ~ "Hmm... It's not a bad idea to visit a doctor.",
    prob >= 20 ~ "You're doing okay. Keep it up!",
    TRUE       ~ "Safe for now — celebrate with an oily pizza and a Diet Coke."
  )

  HTML(glue::glue(
    "<div style='margin-top:20px; font-size:20px; color:#FDF7F7; text-align:center;'>
      {message}<br><span style='font-size:24px; font-weight:bold;'>{prob}%</span>
    </div>"
  ))
})
```

```{r prediction_output}
gaugeOutput("prediction_gauge", height = "300px")
htmlOutput("prediction_text")
```

### Stroke Awareness Video

```{r stroke_video, echo=FALSE}
tags$video(
  src = "stroke.mp4",
  type = "video/mp4",
  controls = NA,
  width = "100%",
  height = "315",
  style = "margin-top: 20px; border-radius: 10px;"
)
```

